Dust trapping in thin-ringed protoplanetary disks

### 2018-08-18 projects/dullemond/paper_dusttraps/ms.tex
Updated dust trapping paper: new radial profiles.

\begin{figure*}
\centerline{\includegraphics[width=0.85\textwidth]{gallery_profiles.pdf}}
\caption{\label{fig-obs-profiles}The linear brightness temperature in
  band 6 of the five disks in our sample which have the most pronounced rings.
  The vertical axis is logarithmic to better show the contrast.
  The eight highest contrast rings are fitted by a Gaussian profile, shown
  as the solid inverse parabolas. The dotted inverse parabolas are Gaussians
  with the width of the ALMA beam. For a detailed description of these
  data, \paperdsharpguzmann{} for AS 209, \paperdsharpisella{}
  for HD 163296, and \paperdsharphuangrings{} for the rest.
{\bf TODO: y-labels}}
\end{figure*}

@@

\begin{figure*}
\centerline{\includegraphics[width=0.85\textwidth]{gaussfits_allrings.pdf}}
\caption{\label{fig-obs-gaussfits}Gaussian fits to the eight rings of this
  paper. The blue curves are the observations, the orange curves are the best fit
  Gaussian profiles. The ``fit range'' bar shows the radial range within which
  the Gauss was fitted to the data. The fit range was chosen to fit the part of
  the curve that, by eye, most resembles a Gaussian. The ``beam'' bar shows the
  FWHM beam size of the observations. The grey band around the blue curve shows
  the estimated uncertainty of the data. {\bf TODO: y-labels}}
\end{figure*}

\begin{figure*}
  \centerline{
    %\includegraphics[width=0.3\textwidth]{AS_209_doublegauss.pdf}
    \includegraphics[width=0.42\textwidth]{HD_163296_doublegauss.pdf}
    \includegraphics[width=0.42\textwidth]{HD_143006_doublegauss.pdf}
  }
  \caption{\label{fig-obs-doublegauss}The sum of the two Gaussian fits
    for the two sources with two partly overlapping rings: HD 163296 and
    HD 143006. {\bf TODO: y-labels}}
\end{figure*}

@@

AS 209     & 1 & 37 &  69 --  79 &   3.64 &   4.18 &  74.1 &  75.4 &   3.87 &   3.38 &  15.8 &  10.9 &   0.6 &   0.48 &  28.5\\
AS 209     & 2 & 37 & 115 -- 125 &   2.95 &   3.23 & 120.3 &  90.3 &   4.64 &   4.23 &  12.4 &   7.7 &   0.4 &   0.55 &  60.7\\
Elias 24   & 1 & 35 &  72 --  82 &   5.22 &   5.69 &  76.9 &  88.0 &   5.08 &   4.66 &  22.3 &  17.2 &   0.6 &   0.40 &  33.4\\
HD 163296  & 1 & 43 &  60 --  76 &   8.48 &   8.76 &  67.4 & 171.7 &   7.36 &   7.13 &  30.8 &  25.6 &   1.6 &   0.42 &  43.8\\
HD 163296  & 2 & 43 &  94 -- 104 &   4.92 &   5.20 &  99.9 & 133.6 &   5.73 &   5.42 &  25.3 &  20.2 &   0.8 &   0.30 &  38.0\\
GW Lup     & 1 & 44 &  79 --  89 &   1.48 &   1.87 &  85.1 &  72.1 &   4.74 &   3.76 &  10.3 &   5.7 &   0.5 &   0.40 &  34.1\\
HD 143006  & 1 & 45 &  35 --  45 &   3.21 &   4.11 &  41.0 &  72.2 &   5.06 &   3.95 &  27.2 &  22.0 &   1.9 &   0.21 &  10.0\\
HD 143006  & 2 & 45 &  59 --  72 &   2.48 &   2.71 &  64.8 & 112.5 &   7.88 &   7.22 &  21.6 &  16.6 &   2.0 &   0.18 &  21.6\\

@@

For large grains and/or weak turbulence one finds $\psi\ll 1$, which leads to
$w_{\mathrm{d}}\ll w$. In this case the dust is strongly trapped near the peak
of the pressure bump. The opposite is the case for small grains and/or strong
turbulence, for which one gets $\psi\gg 1$, which leads to
$w_{\mathrm{d}}\rightarrow w$. In this case the trapping is very weak and the
dust-to-gas ratio within the pressure bump stays constant. Only in the very
wings of the Gaussian pressure bump is dust trapping significant, but this
effect can only be studied using the more accurate non-gaussian solution of
Eq.~(\ref{eq-analytic-sol-radial-trapping-better}). We will return to the dust
trapping effect in the wings in Section {\bf XXXXX}.

Eq.~(\ref{eq-analytic-sol-radial-trapping-summary}) has only two parameters:
$\Sigma_{\mathrm{d0}}$ and $w_{\mathrm{d}}$. Both can be fairly directly
measured from the observations, though $\Sigma_{\mathrm{d0}}$ will depend on the
opacity, and hence on the unknown grain size (see discussion in Appendix
\ref{sec-computing-dust-mass}). The values of $w_{\mathrm{d}}$ for
the rings in our sample can be directly taken from Table \ref{tab-gauss-params}.

@@

One can, however, set an upper bound on $\Sigma_{\mathrm{g}}$ by demanding that
the disk is gravitationally stable, i.e.\ that the Toomre parameter obeys
\begin{equation}
Q_{\mathrm{Toomre}}\equiv \frac{c_s\Omega_K}{\pi G \Sigma_{\mathrm{g}}} >2
\end{equation}
Here $c_s=\sqrt{k_BT/\mu m_p}$ is the isothermal sound speed, with $k_B$ the
Boltzmann constant, $m_p$ the proton mass, and $mu=2.3$ the mean molecular
weight in atomic units. $\Omega_K=\sqrt{GM_{*}/r^3}$ is the Kepler frequency,
$G$ is the gravitational constant and, finally, $\Sigma_{\mathrm{g}}$ the gas
surface density. Taking the disk midplane
temperature from Table \ref{tab-gauss-params}, which was calculated using
Eq.~(\ref{eq-disk-temperature-model}), we can compute the upper limits on
$\Sigma_{\mathrm{g}}$ for all of the rings. The results are given in Table
\ref{tab-ring-model-limits}.

\begin{table*}
\begin{center}
\begin{tabular}{|cc|ccccccc|ccc|}
\hline
\hline
Source     & Ring & $w_{\mathrm{min}}$ & $w_{\mathrm{max}}$ & $\Sigma_{\mathrm{g,min,iv}}$ & $\Sigma_{\mathrm{g,min,ri}}$ & $\Sigma_{\mathrm{g,max}}$ & $a_{\mathrm{max}}$ & $\mathrm{St}_{(a=0.02\,\mathrm{cm})}$ & $\alpha/\mathrm{St}$ & $\alpha/\mathrm{St}$ & $\alpha_{\mathrm{exmp}}$\\
           &      & $[\mathrm{au}]$  & $[\mathrm{au}]$  & $[\mathrm{g}/\mathrm{cm}^3]$ & $[\mathrm{g}/\mathrm{cm}^3]$ & $[\mathrm{g}/\mathrm{cm}^3]$ & $[\mathrm{cm}]$ & (for $\Sigma_{\mathrm{g,max}}$)& (for $w_\mathrm{max}$) & (for $w_\mathrm{min}$) &   \\
\hline
% Note: New table made with new stellar parameters 2018.07.02
% Note: New table with new beam size estimates 2018.08.06
AS 209     & 1 &  5.59 & 19.6 & 1.1e-02 & 2.0e-01 & 1.6e+01 & 15.7 & 3.9e-03 & 3.1e-02 & 5.7e-01 & 1.2e-04\\
AS 209     & 2 & 10.25 & 19.6 & 1.2e-02 & 2.3e-01 & 6.9e+00 &  5.8 & 9.1e-03 & 4.9e-02 & 2.1e-01 & 4.4e-04\\
Elias 24   & 1 &  7.26 & 17.1 & 9.7e-03 & 1.8e-01 & 1.8e+01 & 18.8 & 3.6e-03 & 8.0e-02 & 7.0e-01 & 2.9e-04\\
HD 163296  & 1 &  7.13$^{*}$ & 13.8 & 1.1e-02 & 1.9e-01 & 4.1e+01 & 40.0 & 1.5e-03 & 3.6e-01 & -- & 5.6e-04\\
HD 163296  & 2 &  7.07 & 13.8 & 7.8e-03 & 1.4e-01 & 2.0e+01 & 27.3 & 3.1e-03 & 1.8e-01 & 1.4e+00 & 5.6e-04\\
GW Lup     & 1 &  7.49 &  9.9 & 8.3e-03 & 1.5e-01 & 7.8e+00 &  9.8 & 8.0e-03 & 1.7e-01 & 3.4e-01 & 1.3e-03\\
HD 143006  & 1 &  3.95$^{*}$ & 10.1 & 4.7e-03 & 8.6e-02 & 7.5e+01 & 167.1 & 8.4e-04 & 1.8e-01 & -- & 1.5e-04\\
HD 143006  & 2 &  7.22$^{*}$ & 10.1 & 4.8e-03 & 8.8e-02 & 3.4e+01 & 73.1 & 1.9e-03 & 1.0e+00 & -- & 1.9e-03\\
\hline
\hline
\end{tabular}
\end{center}
\caption{\label{tab-ring-model-limits}Limits on the free parameters of the dust
  trapping model. The lower limit to the pressure bump width $w_{\mathrm{min}}$
  is the pressure scale height $h_p$. If, however, the width of the dust ring
  $w_d>h_p$, then the lower limit is $w_d$ (marked with the symbol $^{*}$).  The
  upper limit $w_{\mathrm{max}}$ is derived from the separation between the
  rings (for AS 209, HD 163296 and HD 143006) or from the separation of the ring
  to the nearest minimum (for Elias 24 and GW Lup). The two lower limits on the
  gas surface density $\Sigma_{\mathrm{g}}$ were derived by demanding
  $\Sigma_{\mathrm{g}}\gtrsim \Sigma_{\mathrm{d}}$, where $\Sigma_{\mathrm{d}}$
  was computed in Appendix \ref{sec-computing-dust-mass} using two different
  opacities: that of Ivezic et al. (leading to $\Sigma_{\mathrm{g,min,iv}}$) and
  that of Ricci et al.\ (leading to $\Sigma_{\mathrm{g,min,ri}}$). The column
  marked with $\Sigma_{\mathrm{g,max}}$ is the upper limit on the gas surface
  density derived from demanding that the gas disk is gravitationally
  stable. The $a_{\mathrm{max}}$ column is the maximum grain size for which the
  derived dust surface density together with the gas surface density remain
  gravitationally stable. The $\mathrm{St}_{(a=0.02\,\mathrm{cm})}$ column gives
  an example value of the Stokes number: it is the value of $\mathrm{St}$ if the
  grains have a radius of $0.02\,\mathrm{cm}$, for the case of
  $\Sigma_{\mathrm{g}}=\Sigma_{\mathrm{g,max}}$. The two columns for
  $\alpha/\mathrm{St}$ are derived for the widest and narrowest
  gas bump, respectively. The column $\alpha_{\mathrm{exmp}}$ is the value of
  $\alpha_{\mathrm{turb}}$ for an example choice of parameters: $w=w_{\mathrm{max}}$,
  $\Sigma_{\mathrm{g}}=\Sigma_{\mathrm{g,max}}$ and $a=0.02\,\mathrm{cm}$.}
\end{table*}

One can estimate a lower limit to the gas density by demanding that the gas
surface density must be at least as large as the dust surface density, since
dust trapping is unlikely to achieve a larger concentration of dust than
that. Depending on which of the two opacities we use (see Appendix
\ref{sec-computing-dust-mass}) we arrive at different estimates of
$\Sigma_{\mathrm{d}}(r=r_0)$. By demanding $\Sigma_{\mathrm{g}}\gtrsim
\Sigma_{\mathrm{d}}$ we arrive at two lower limits of the gas surface density, given in
Table \ref{tab-ring-model-limits}. It is likely that even for much
larger values of the gas surface density the dust-gas mixture becomes
unstable to the streaming instability and other types of instabilities.
If we can convince ourselves that such instabilities are not operating
in these rings, then this puts a substantially stronger (i.e.~larger)
lower limit on the gas surface density.

@@

For the choice $w=w_{\mathrm{max}}$ (the widest possible pressure bump) the dust
rings are all narrower that the gas rings: $w_{\mathrm{d}}<w$, which implies
that the dust trapping is operational, and Eq.~(\ref{eq-alphaSt-equation}) gives
information about the turbulent strength. For the choice $w=w_{\mathrm{min}}$
(the narrowest possible pressure bump) we can only use
Eq.~(\ref{eq-alphaSt-equation}) for the rings for which
$w_{\mathrm{d}}<h_p$. The reason is that for those rings with
$w_{\mathrm{d}}>h_p$ (marked with a $^{*}$ in Table \ref{tab-ring-model-limits})
the minimal pressure bump width is $w_{\mathrm{min}}=w_{\mathrm{d}}$, and the
dust ring is as wide as the pressure bump, implying that dust trapping is weak
or non-operational. Any increase of $\alpha_{\mathrm{turb}}/\mathrm{St}$ will
keep $w_{\mathrm{d}}=w_{\mathrm{min}}$, so one cannot derive any value for
$\alpha_{\mathrm{turb}}/\mathrm{St}$. But for other rings we can compute
$\alpha_{\mathrm{turb}}/\mathrm{St}$. The resulting values for both choices of
pressure bump width are given in Table \ref{tab-ring-model-limits}. They can be
understood as the lower and upper limit on $\alpha_{\mathrm{turb}}/\mathrm{St}$.

@@

These low values of $\alpha_{\mathrm{turb}}$ are consistent with the low values
or upper limits reported recently \citep{2016ApJ...816...25P,
  2018ApJ...856..117F} {\bf [more?]}. However, it has to be kept in mind that
the values of $\alpha_{\mathrm{turb}}=\alpha_{\mathrm{exmp}}$ were derived
for an example choice of parameters: : $w=w_{\mathrm{max}}$,
$\Sigma_{\mathrm{g}}=\Sigma_{\mathrm{g,max}}$ and $a=0.02\,\mathrm{cm}$. For a
smaller value of $w$, a lower value of $\Sigma_{\mathrm{g}}$, or larger grains,
the computed value of $\alpha_{\mathrm{turb}}$ will increase. So that it is hard
to set a true upper limit on $\alpha_{\mathrm{turb}}$ from these observations.

@@

Should the low spectral slope at the peak of the rings be confirmed, then the
lower limits on $\alpha_{\mathrm{turb}}$ from the $\alpha_{\mathrm{exmp}}$
column in Table \ref{tab-ring-model-limits} paint a different picture from the
low-turbulence picture that has emerged in recent times. The values of a few
times $10^{-4}$ to a few times $10^{-3}$ are low, but they are only lower
bounds. It would be somewhat unlikely if all rings have parameters at this
extreme size of parameter space, and it is therefore likely that
$\alpha_{\mathrm{turb}}$ is easily a factor of ten higher or so. This would
imply that protoplanetary disks are substantially turbulent after all. But to
confirm this, we need to dig a bit deeper by including size distributions
(Section \ref{sec-model-with-grain-size-distribution}) and non-perfect
(``leaky'') dust traps (Section {\bf XXXX}).

{\bf We can make predictions for pressure slopes in CO}

{\bf CONSIDER PRESSURE OFFSET!}

@@

\begin{acknowledgements}
  The authors acknowledge support
  by the High Performance and Cloud Computing Group at the Zentrum f\"ur
  Datenverarbeitung of the University of T\"ubingen, the state of
  Baden-W\"urttemberg through bwHPC and the German Research Foundation (DFG)
  through grant no INST 37/935-1 FUGG. This research was initiated at the
  ``Stars, Planets and Galaxies'' meeting at the Harnackhaus in Berlin,
  April 2018. Part of this work was also funded by the DFG Forschergruppe
  FOR 2634 ``Planet Formation Witnesses and Probes: Transition Disks''.
\end{acknowledgements}

@@

\begin{table}
\begin{center}
\begin{tabular}{|cc|ccc|}
\hline
\hline
Source & Ring & $A$ & $r_0$ & $\sigma$ \\
\hline
AS 209     & 1 & $  3.637_{-0.047}^{+0.048}$ & $ 74.129_{-0.066}^{+0.063}$ & $  3.885_{-0.090}^{+0.098}$  \\
AS 209     & 2 & $  2.949_{-0.031}^{+0.030}$ & $120.309_{-0.068}^{+0.065}$ & $  4.651_{-0.111}^{+0.115}$  \\
Elias 24   & 1 & $  5.214_{-0.068}^{+0.066}$ & $ 76.851_{-0.092}^{+0.088}$ & $  5.104_{-0.172}^{+0.186}$  \\
HD 163296  & 1 & $  8.475_{-0.056}^{+0.056}$ & $ 67.403_{-0.055}^{+0.057}$ & $  7.370_{-0.098}^{+0.098}$  \\
HD 163296  & 2 & $  4.918_{-0.026}^{+0.027}$ & $ 99.862_{-0.061}^{+0.059}$ & $  5.741_{-0.116}^{+0.119}$  \\
GW Lup     & 1 & $  1.467_{-0.052}^{+0.055}$ & $ 85.178_{-0.259}^{+0.316}$ & $  4.882_{-0.402}^{+0.530}$  \\
HD 143006  & 1 & $  3.202_{-0.084}^{+0.082}$ & $ 41.080_{-0.169}^{+0.190}$ & $  5.133_{-0.320}^{+0.373}$  \\
HD 143006  & 2 & $  2.473_{-0.043}^{+0.044}$ & $ 64.831_{-0.248}^{+0.230}$ & $  8.029_{-0.501}^{+0.593}$  \\
\hline
\hline
\end{tabular}
\end{center}
\caption{\label{tab-gauss-fit-errors}The Gaussian fit values with their error
  estimates from the MCMC procedure described in Appendix \ref{sec-gauss-fitting-procedure}.}
\end{table}

>>>>

\begin{figure*}
\centerline{\includegraphics[width=0.85\textwidth]{gallery_profiles.pdf}}
\caption{\label{fig-obs-profiles}The linear brightness temperature in
  band 6 of the five disks in our sample which have the most pronounced rings.
  The vertical axis is logarithmic to better show the contrast.
  The eight highest contrast rings are fitted by a Gaussian profile, shown
  as the solid inverse parabolas. The dotted inverse parabolas are Gaussians
  with the width of the ALMA beam. For a detailed description of these
  data, see \paperdsharphuangrings{}.}
\end{figure*}

@@

\begin{figure*}
\centerline{\includegraphics[width=0.85\textwidth]{gaussfits_allrings.pdf}}
\caption{\label{fig-obs-gaussfits}Gaussian fits to the eight rings of this
  paper. The blue curves are the observations, the orange curves are the best fit
  Gaussian profiles. The ``fit range'' bar shows the radial range within which
  the Gauss was fitted to the data. The fit range was chosen to fit the part of
  the curve that, by eye, most resembles a Gaussian. The ``beam'' bar shows the
  FWHM beam size of the observations. The grey band around the blue curve shows
  the estimated uncertainty of the data.}
\end{figure*}

\begin{figure*}
  \centerline{
    %\includegraphics[width=0.3\textwidth]{AS_209_doublegauss.pdf}
    \includegraphics[width=0.42\textwidth]{HD_163296_doublegauss.pdf}
    \includegraphics[width=0.42\textwidth]{HD_143006_doublegauss.pdf}
  }
  \caption{\label{fig-obs-doublegauss}The sum of the two Gaussian fits
    for the two sources with two partly overlapping rings: HD 163296 and
    HD 143006.}
\end{figure*}

@@

AS 209     & 1 & 37 &  69 --  79 &   3.66 &   4.17 &  74.1 &  77.6 &   3.99 &   3.51 &  15.8 &  10.9 &   0.6 &   0.48 &  29.2\\
AS 209     & 2 & 37 & 115 -- 125 &   2.98 &   3.26 & 120.4 &  91.9 &   4.72 &   4.32 &  12.4 &   7.7 &   0.4 &   0.55 &  62.3\\
Elias 24   & 1 & 35 &  72 --  82 &   5.50 &   6.03 &  76.7 &  85.3 &   4.92 &   4.49 &  22.3 &  17.2 &   0.6 &   0.43 &  34.1\\
HD 163296  & 1 & 43 &  60 --  76 &   8.75 &   9.08 &  67.7 & 161.4 &   6.92 &   6.67 &  30.8 &  25.6 &   1.5 &   0.44 &  43.0\\
HD 163296  & 2 & 43 &  94 -- 104 &   5.18 &   5.54 & 100.0 & 120.3 &   5.16 &   4.82 &  25.3 &  20.2 &   0.7 &   0.32 &  36.6\\
GW Lup     & 1 & 44 &  79 --  89 &   1.29 &   1.49 &  85.5 &  87.7 &   5.77 &   4.99 &  10.2 &   5.7 &   0.7 &   0.30 &  33.5\\
HD 143006  & 1 & 45 &  35 --  45 &   3.31 &   4.22 &  41.0 &  72.6 &   5.09 &   3.99 &  27.2 &  22.0 &   1.9 &   0.21 &  10.3\\
HD 143006  & 2 & 45 &  59 --  72 &   2.58 &   2.80 &  65.2 & 114.3 &   8.01 &   7.37 &  21.6 &  16.5 &   2.0 &   0.19 &  22.9\\

@@

For large grains and/or weak turbulence one finds $\psi\ll 1$, which leads to
$w_{\mathrm{d}}\ll w$. In this case the dust is strongly trapped near the peak
of the pressure bump. The opposite is the case for small grains and/or strong
turbulence, for which one gets $\psi\gg 1$, which leads to
$w_{\mathrm{d}}\rightarrow w$. In this case the trapping is very weak and the
dust-to-gas ratio within the pressure bump stays constant. Only in the very
wings of the Gaussian pressure bump is dust trapping then significant, but this
effect can only be studied using the more accurate non-gaussian solution of
Eq.~(\ref{eq-analytic-sol-radial-trapping-better}). We will return to the dust
trapping effect in the wings in Section {\bf XXXXX}.

Eq.~(\ref{eq-analytic-sol-radial-trapping-summary}) has only two parameters:
$\Sigma_{\mathrm{d0}}$ and $w_{\mathrm{d}}$. As we have shown in Sections
\ref{sec-gauss-fits}, \ref{sec-opt-thin-analysis} and
\ref{sec-optical-depth-effects}, both parameter can be extracted from the
observations. The main uncertainty lies in $\Sigma_{\mathrm{d0}}$, due to the
uncertainty in the dust opacity. The values of $w_{\mathrm{d}}$ for the rings in
our sample can be directly taken from Table \ref{tab-gauss-params}.

@@

One can, however, set an upper bound on $\Sigma_{\mathrm{g}}$ by demanding that
the disk is gravitationally stable, i.e.\ that the Toomre parameter obeys
\begin{equation}
Q_{\mathrm{Toomre}}\equiv \frac{c_s\Omega_K}{\pi G \Sigma_{\mathrm{g}}} >2
\end{equation}
Here $c_s=\sqrt{k_BT/\mu m_p}$ is the isothermal sound speed. Here
$\Omega_K=\sqrt{GM_{*}/r^3}$ is the Kepler frequency,
$G$ is the gravitational constant and $\Sigma_{\mathrm{g}}$ the gas
surface density. Taking the disk midplane
temperature from Table \ref{tab-gauss-params}, which was calculated using
Eq.~(\ref{eq-disk-temperature-model}), we can compute the upper limits on
$\Sigma_{\mathrm{g}}$ for all of the rings. The results are given in Table
\ref{tab-ring-model-limits}.

\begin{table*}
\begin{center}
\begin{tabular}{|cc|ccccccc|ccc|}
\hline
\hline
Source     & Ring & $w_{\mathrm{min}}$ & $w_{\mathrm{max}}$ & $\Sigma_{\mathrm{g,min,iv}}$ & $\Sigma_{\mathrm{g,min,ri}}$ & $\Sigma_{\mathrm{g,max}}$ & $a_{\mathrm{max}}$ & $\mathrm{St}_{(a=0.02\,\mathrm{cm})}$ & $\alpha/\mathrm{St}$ & $\alpha/\mathrm{St}$ & $\alpha_{\mathrm{exmp}}$\\
           &      & $[\mathrm{au}]$  & $[\mathrm{au}]$  & $[\mathrm{g}/\mathrm{cm}^3]$ & $[\mathrm{g}/\mathrm{cm}^3]$ & $[\mathrm{g}/\mathrm{cm}^3]$ & $[\mathrm{cm}]$ & (for $\Sigma_{\mathrm{g,max}}$)& (for $w_\mathrm{max}$) & (for $w_\mathrm{min}$) &   \\
\hline
% Note: New table made with new stellar parameters 2018.07.02
% Note: New table with new beam size estimates 2018.08.06
AS 209     & 1 &  5.59 & 19.6 & 1.1e-01 & 1.7e-01 & 1.6e+01 &  1.5 & 3.9e-03 & 3.3e-02 & 6.5e-01 & 1.3e-04\\
AS 209     & 2 & 10.26 & 19.6 & 1.2e-01 & 1.9e-01 & 6.9e+00 &  0.6 & 9.1e-03 & 5.1e-02 & 2.2e-01 & 4.6e-04\\
Elias 24   & 1 &  7.24 & 17.1 & 1.0e-01 & 1.5e-01 & 1.8e+01 &  1.8 & 3.6e-03 & 7.4e-02 & 6.3e-01 & 2.6e-04\\
HD 163296  & 1 &  6.67$^{*}$ & 13.8 & 1.0e-01 & 1.6e-01 & 4.0e+01 &  4.1 & 1.6e-03 & 3.1e-01 & -- & 4.7e-04\\
HD 163296  & 2 &  7.08 & 13.8 & 8.0e-02 & 1.2e-01 & 2.0e+01 &  2.7 & 3.1e-03 & 1.4e-01 & 8.6e-01 & 4.3e-04\\
GW Lup     & 1 &  7.53 &  9.9 & 7.6e-02 & 1.2e-01 & 7.8e+00 &  1.1 & 8.1e-03 & 3.4e-01 & 7.8e-01 & 2.8e-03\\
HD 143006  & 1 &  3.99$^{*}$ & 10.1 & 5.6e-02 & 8.4e-02 & 7.5e+01 & 14.0 & 8.3e-04 & 1.9e-01 & -- & 1.5e-04\\
HD 143006  & 2 &  7.37$^{*}$ & 10.1 & 4.9e-02 & 7.4e-02 & 3.4e+01 &  7.0 & 1.9e-03 & 1.1e+00 & -- & 2.1e-03\\
\hline
\hline
\end{tabular}
\end{center}
\caption{\label{tab-ring-model-limits}Limits on the free parameters of the dust
  trapping model. The lower limit to the pressure bump width $w_{\mathrm{min}}$
  is the pressure scale height $h_p$. If, however, the width of the dust ring
  $w_d>h_p$, then the lower limit is $w_d$ (marked with the symbol $^{*}$).  The
  upper limit $w_{\mathrm{max}}$ is derived from the separation between the
  rings (for AS 209, HD 163296 and HD 143006) or from the separation of the ring
  to the nearest minimum (for Elias 24 and GW Lup). The two lower limits on the
  gas surface density $\Sigma_{\mathrm{g}}$ were derived by demanding
  $\Sigma_{\mathrm{g}}\gtrsim \Sigma_{\mathrm{d}}$, where $\Sigma_{\mathrm{d}}$
  was computed in Appendix \ref{sec-computing-dust-mass} using two different
  opacities: that of Ivezic et al. (leading to $\Sigma_{\mathrm{g,min,iv}}$) and
  that of Ricci et al.\ (leading to $\Sigma_{\mathrm{g,min,ri}}$). The column
  marked with $\Sigma_{\mathrm{g,max}}$ is the upper limit on the gas surface
  density derived from demanding that the gas disk is gravitationally
  stable. The $a_{\mathrm{max}}$ column is the maximum grain size for which the
  derived dust surface density together with the gas surface density remain
  gravitationally stable. The $\mathrm{St}_{(a=0.02\,\mathrm{cm})}$ column gives
  an example value of the Stokes number: it is the value of $\mathrm{St}$ if the
  grains have a radius of $0.02\,\mathrm{cm}$, for the case of
  $\Sigma_{\mathrm{g}}=\Sigma_{\mathrm{g,max}}$. The two columns for
  $\alpha/\mathrm{St}$ are derived for the widest and narrowest
  gas bump, respectively. The column $\alpha_{\mathrm{exmp}}$ is the value of
  $\alpha_{\mathrm{turb}}$ for an example choice of parameters: $w=w_{\mathrm{max}}$,
  $\Sigma_{\mathrm{g}}=\Sigma_{\mathrm{g,max}}$ and $a=0.02\,\mathrm{cm}$.}
\end{table*}

One can estimate a lower limit to the gas density by demanding that the gas
surface density must be at least as large as the dust surface density, since
dust trapping is unlikely to achieve a larger concentration of dust than
that. Depending on which of the two opacities we use (see Appendix
\ref{sec-computing-dust-mass}) {\bf TODO: OPACITIES AND IMPLICATIONS} we arrive at different estimates of
$\Sigma_{\mathrm{d}}(r=r_0)$. By demanding $\Sigma_{\mathrm{g}}\gtrsim
\Sigma_{\mathrm{d}}$ we arrive at two lower limits of the gas surface density, given in
Table \ref{tab-ring-model-limits}. It is likely that even for much
larger values of the gas surface density the dust-gas mixture becomes
unstable to the streaming instability and other types of instabilities.
If we can convince ourselves that such instabilities are not operating
in these rings, then this puts a substantially stronger (i.e.~larger)
lower limit on the gas surface density.

@@

For the choice $w=w_{\mathrm{max}}$ (the widest possible pressure bump) the dust
rings are all narrower that the gas rings: $w_{\mathrm{d}}<w$, which implies
that the dust trapping is operational, and Eq.~(\ref{eq-alphaSt-equation}) gives
information about the turbulent strength. For the choice $w=w_{\mathrm{min}}$
(the narrowest possible pressure bump) we can only use
Eq.~(\ref{eq-alphaSt-equation}) for the rings for which
$w_{\mathrm{d}}<h_p$. The reason is that for those rings with
$w_{\mathrm{d}}>h_p$ (marked with a $^{*}$ in Table \ref{tab-ring-model-limits})
the minimal pressure bump width is $w_{\mathrm{min}}=w_{\mathrm{d}}$, and the
dust ring is as wide as the pressure bump, implying that dust trapping is weak
or non-operational. Any increase of $\alpha_{\mathrm{turb}}/\mathrm{St}$ will
keep $w_{\mathrm{d}}=w_{\mathrm{min}}$, so one cannot derive any value for
$\alpha_{\mathrm{turb}}/\mathrm{St}$. But for other rings (those not marked
with *) we can compute
$\alpha_{\mathrm{turb}}/\mathrm{St}$. The resulting values for both choices of
pressure bump width are given in Table \ref{tab-ring-model-limits}. They can be
understood as the lower and upper limit on $\alpha_{\mathrm{turb}}/\mathrm{St}$.

@@

These low values of $\alpha_{\mathrm{turb}}$ are consistent with the low values
or upper limits reported recently \citep{2016ApJ...816...25P,
  2018ApJ...856..117F} {\bf [more?]}. However, it has to be kept in mind that
the values of $\alpha_{\mathrm{turb}}=\alpha_{\mathrm{exmp}}$ were derived for
an example choice of parameters: : $w=w_{\mathrm{max}}$,
$\Sigma_{\mathrm{g}}=\Sigma_{\mathrm{g,max}}$ and $a=0.02\,\mathrm{cm}$. For a
smaller value of $w$, a lower value of $\Sigma_{\mathrm{g}}$, or larger grains,
the computed value of $\alpha_{\mathrm{turb}}$ will increase. This means that it
is hard to set a true upper limit on $\alpha_{\mathrm{turb}}$ from these
observations.

@@

Should the low spectral slope at the peak of the rings be confirmed, then the
lower limits on $\alpha_{\mathrm{turb}}$ from the $\alpha_{\mathrm{exmp}}$
column in Table \ref{tab-ring-model-limits} paint a different picture from the
low-turbulence picture that has emerged in recent times. The values of a few
times $10^{-4}$ to a few times $10^{-3}$ are low, but they are only lower
bounds. It would be somewhat unlikely if all rings have parameters at this
extreme size of parameter space, and it is therefore likely that
$\alpha_{\mathrm{turb}}$ is easily a factor of ten higher or so. This would
imply that protoplanetary disks are substantially turbulent after all. But to
confirm this, we need to dig a bit deeper by including size distributions
(Section \ref{sec-model-with-grain-size-distribution}).

\subsection{Non-perfect trapping: leakage of dust}
\label{sec-leaky-traps}
%
So far we have assumed that the Gaussian pressure bump remains perfectly
Gaussian well into the wings. This is obviously an unrealistic assumption. In
practice, the bump will level off into a background pressure profile
$p_{\mathrm{bg}}(r)$ that varies much slower with $r$ than the bump. In its
simplest form we can investigate the following case:
\begin{equation}
  p(r) = \mathrm{max}\left(
  p_0 \exp\left(-\frac{(r-r_0)^2}{2w^2}\right),
  p_{\mathrm{bg}}(r)\right)
\end{equation}
which simply switches between the Gaussian bump of
Eq.~(\ref{eq-gaussian-pressure-bump}) and the background profile
$p_{\mathrm{bg}}(r)$, depending on which of the two is higher. Usually the
background pressure profile has a negative slope: $dp_{\mathrm{bg}}(r)/dr<0$, so
that any dust that reaches .. {\bf [NOT SURE IF THIS IS CORRECT]}

To estimate
the time it takes for the trapped dust to leak out

@@

\begin{acknowledgements}
  The authors acknowledge support by the High Performance and Cloud Computing
  Group at the Zentrum f\"ur Datenverarbeitung of the University of T\"ubingen,
  the state of Baden-W\"urttemberg through bwHPC and the German Research
  Foundation (DFG) through grant no INST 37/935-1 FUGG. Part of this work was
  also funded by the DFG Forschergruppe FOR 2634 ``Planet Formation Witnesses
  and Probes: Transition Disks''.
\end{acknowledgements}

@@

\begin{table}
\begin{center}
\begin{tabular}{|cc|ccc|}
\hline
\hline
Source & Ring & $A$ & $r_0$ & $\sigma$ \\
\hline
AS 209     & 1 & $  3.663_{-0.045}^{+0.048}$ & $ 74.134_{-0.067}^{+0.069}$ & $  3.989_{-0.097}^{+0.106}$  \\
AS 209     & 2 & $  2.981_{-0.031}^{+0.031}$ & $120.392_{-0.074}^{+0.073}$ & $  4.722_{-0.117}^{+0.129}$  \\
Elias 24   & 1 & $  5.498_{-0.061}^{+0.060}$ & $ 76.650_{-0.072}^{+0.073}$ & $  4.923_{-0.122}^{+0.138}$  \\
HD 163296  & 1 & $  8.749_{-0.073}^{+0.079}$ & $ 67.683_{-0.069}^{+0.071}$ & $  6.922_{-0.116}^{+0.120}$  \\
HD 163296  & 2 & $  5.178_{-0.035}^{+0.033}$ & $ 99.963_{-0.059}^{+0.066}$ & $  5.161_{-0.114}^{+0.117}$  \\
GW Lup     & 1 & $  1.292_{-0.036}^{+0.034}$ & $ 85.526_{-0.295}^{+0.399}$ & $  5.771_{-0.471}^{+0.597}$  \\
HD 143006  & 1 & $  3.312_{-0.073}^{+0.075}$ & $ 40.996_{-0.193}^{+0.228}$ & $  5.087_{-0.319}^{+0.371}$  \\
HD 143006  & 2 & $  2.577_{-0.051}^{+0.056}$ & $ 65.161_{-0.267}^{+0.243}$ & $  8.011_{-0.544}^{+0.672}$  \\
\hline
\hline
\end{tabular}
\end{center}
\caption{\label{tab-gauss-fit-errors}The Gaussian fit values with their error
  estimates from the MCMC procedure described in Appendix \ref{sec-gauss-fitting-procedure}.}
\end{table}
